---
title: "GEO880 - CMA: Exercise 2"
author: "Gregory Biland"
date: "5/3/2022"
output:
  html_document:
    code_folding: hide
    fig_caption: yes
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: no
  pdf_document:
    toc: yes
    toc_depth: '3'
    fig_caption: yes
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
Sys.setenv(LANG = "en")
library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times
library(grid)
library(gridExtra)
library(zoo)

dataFolder   <- here::here()   
RFolder      <- here::here()         
figureFolder <- here::here("figs") 
```

## Taks 1
```{r,echo = FALSE, warning = FALSE, message = FALSE}
wildschwein_BE <- read_delim(file.path(dataFolder, "wildschwein_BE_2056.csv")) # adjust path

```

```{r,echo = FALSE, warning = FALSE, message = FALSE}
wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)

```

## Task 2

```{r,echo = FALSE, warning = FALSE, message = FALSE}
wildschwein <- wildschwein_BE %>% group_by(TierID) %>% mutate(timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC)))

wildschwein

wildschwein %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_timelag = mean(timelag,na.rm = T))

summary(wildschwein$timelag)

```
For how long were the individual tracked? Are there gaps?
Were all individuals tracked concurrently or sequentially?
What is the temporal sampling interval between the locations?

```{r,echo = FALSE, warning = FALSE, message = FALSE}
ggplot()+
  geom_line(data = wildschwein, aes(x = DatetimeUTC, y = TierID))

ggplot()+
  geom_step(data = wildschwein, aes(x = DatetimeUTC, y = timelag, col = TierID))
```
In total 3 individuals were tracked, corresponding to the numbers "002A", "016A" and "018A". All animals, were tracked at the same time but not with the same duration. The animal with the number 002A, was monitored the longest from fall of 2014 until the summer of 2015. Regarding the gaps in the monitoring, I would not argue that there were gaps, despite the fact that the largest time gap between to measurements was almost 1 day, where the average timelag was only 23 Minutes.


## Task 3

```{r,echo = FALSE, warning = FALSE, message = FALSE}
wildschwein <- wildschwein %>% group_by(TierID) %>% 
    mutate(steplength = sqrt((E - lead(E,1))^2+ (N - lead(N,1))^2))

wildschwein <- wildschwein %>% mutate(speed = (steplength / timelag)*3.6)

ggplot() +
  geom_line(data = wildschwein, aes(y = speed, x = DatetimeUTC, col = TierID))+
  xlab("Time")+
  ylab("Speed [km/h]")+
  ggtitle("Speed over time for each of the three observed animals [002A, 0016A, 0018A]")

wildschwein %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_speed = mean(speed,na.rm = T))

```
The unit in which the speed is returned would be m/s, but can be changed to km/h, by multiplying by 3.6.

## Task 4

```{r,echo = FALSE, warning = FALSE, message = FALSE}
caro <- read_delim(file.path(dataFolder, "caro60.csv"))
```

```{r,echo = FALSE, warning = FALSE, message = FALSE}
caro3 <- caro %>% slice(seq(1,200, by = 3))
caro6 <- caro %>% slice(seq(1,200, by = 6))
caro9 <- caro %>% slice(seq(1,200, by = 9))
```

```{r Cora,echo = FALSE, warning = FALSE, message = FALSE}
caro <- caro %>% group_by(TierID) %>% mutate(timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC))*60)

caro %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_timelag = mean(timelag,na.rm = T))

caro <- caro %>% group_by(TierID) %>% 
    mutate(steplength = sqrt((E - lead(E,1))^2+ (N - lead(N,1))^2))

caro <- caro %>% mutate(speed = (steplength / timelag)*3.6)

caro %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_speed = mean(speed,na.rm = T))

```

```{r Caro3,echo = FALSE, warning = FALSE, message = FALSE}
caro3 <- caro3 %>% group_by(TierID) %>% mutate(timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC))*60)

caro3 %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_timelag = mean(timelag,na.rm = T))

caro3 <- caro3 %>% group_by(TierID) %>% 
    mutate(steplength = sqrt((E - lead(E,1))^2+ (N - lead(N,1))^2))

caro3 <- caro3 %>% mutate(speed = (steplength / timelag)*3.6)

caro3 %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_speed = mean(speed,na.rm = T))

```

```{r Caro6,echo = FALSE, warning = FALSE, message = FALSE}
caro6 <- caro6 %>% group_by(TierID) %>% mutate(timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC))*60)

caro6 %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_timelag = mean(timelag,na.rm = T))

caro6 <- caro6 %>% group_by(TierID) %>% 
    mutate(steplength = sqrt((E - lead(E,1))^2+ (N - lead(N,1))^2))

caro6 <- caro6 %>% mutate(speed = (steplength / timelag)*3.6)

caro6 %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_speed = mean(speed,na.rm = T))
```

```{r Caro9,echo = FALSE, warning = FALSE, message = FALSE}
caro9 <- caro9 %>% group_by(TierID) %>% mutate(timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC))*60)

caro9 %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_timelag = mean(timelag,na.rm = T))

caro9 <- caro9 %>% group_by(TierID) %>% 
    mutate(steplength = sqrt((E - lead(E,1))^2+ (N - lead(N,1))^2))

caro9 <- caro9 %>% mutate(speed = (steplength / timelag)*3.6)

caro9 %>%                           
  group_by(TierID) %>%                    
  summarise(                              
    mean_speed = mean(speed,na.rm = T))
```

```{r,echo = FALSE, warning = FALSE, message = FALSE}
c1 <- ggplot() +
  geom_line(data = caro, aes(y = speed, x = DatetimeUTC, col = TierID))+
  xlab("Time")+
  ylab("Speed [km/h]")

c3 <- ggplot() +
  geom_line(data = caro3, aes(y = speed, x = DatetimeUTC, col = TierID))+
  xlab("Time")+
  ylab("Speed [km/h]")

c6 <- ggplot() +
  geom_line(data = caro6, aes(y = speed, x = DatetimeUTC, col = TierID))+
  xlab("Time")+
  ylab("Speed [km/h]")

c9 <- ggplot() +
  geom_line(data = caro9, aes(y = speed, x = DatetimeUTC, col = TierID))+
  xlab("Time")+
  ylab("Speed [km/h]")

grid.arrange(c1, c3, c6, c9, nrow = 2, top = textGrob("Comparing derived speed at different sampling intervals",gp=gpar(fontsize=20,font=3)))


ggplot()+
  geom_path(data = caro, aes(x = E, y = N, col = "red"), alpha = 0.3)+
  geom_point(data = caro, aes(x = E, y = N, col = "red"), alpha = 0.3)+
  geom_path(data = caro3, aes(x = E, y = N, col = "blue"))+
  geom_point(data = caro3, aes(x = E, y = N, col = "blue"))+
  ggtitle("Comparing original- with 3 minutes-resampled data")+ labs(colour = "Trajectory")+
  theme_minimal()+
  scale_color_manual(labels = c("1- minute", "3- minutes"),
                     values = c("red", "blue"))

ggplot()+
  geom_path(data = caro, aes(x = E, y = N, col = "red"), alpha = 0.3)+
  geom_point(data = caro, aes(x = E, y = N, col = "red"), alpha = 0.3)+
  geom_path(data = caro6, aes(x = E, y = N, col = "blue"))+
  geom_point(data = caro6, aes(x = E, y = N, col = "blue"))+
  ggtitle("Comparing original- with 6 minutes-resampled data")+
  labs(colour = "Trajectory")+
  theme_minimal()+
  scale_color_manual(labels = c("1- minute", "6- minutes"),
                     values = c("red", "blue"))

ggplot()+
  geom_path(data = caro, aes(x = E, y = N, col = "red"), alpha = 0.3)+
  geom_point(data = caro, aes(x = E, y = N, col = "red"), alpha = 0.3)+
  geom_path(data = caro9, aes(x = E, y = N, col = "blue"))+
  geom_point(data = caro9, aes(x = E, y = N, col = "blue"))+
  ggtitle("Comparing original- with 9 minutes-resampled data")+
  labs(colour = "Trajectory")+
  theme_minimal()+
  scale_color_manual(labels = c("1- minute", "9- minutes"),
                     values = c("red", "blue"))

```

## Task 5

```{r,echo = FALSE, warning = FALSE, message = FALSE}
caro_rw3 <- rollmean(caro$speed,k = 3,fill = NA,align = "left") %>% as_data_frame()
summary(caro_rw3)
caro_rw4 <- rollmean(caro$speed,k = 4,fill = NA,align = "left") %>% as_data_frame()
summary(caro_rw4)
caro_rw5 <- rollmean(caro$speed,k = 5,fill = NA,align = "left") %>% as_data_frame()
summary(caro_rw5)
caro_rw6 <- rollmean(caro$speed,k = 6,fill = NA,align = "left") %>% as_data_frame()
summary(caro_rw6)


ggplot()+
  geom_line(data = caro_rw3, aes(y = value, x = caro$DatetimeUTC,col = "red"))+
  geom_line(data = caro_rw4, aes(y = value, x = caro$DatetimeUTC,col = "green"))+
  geom_line(data = caro_rw5, aes(y = value, x = caro$DatetimeUTC,col = "blue"))+
  geom_line(data = caro_rw6, aes(y = value, x = caro$DatetimeUTC,col = "darkred"))+
  ggtitle("Comparing different k- values in moving window function - rollmean")+
  labs(colour = "k- values")+
  theme_minimal()+
  xlab("Time")+
  ylab("Speed [km/h]")+
  scale_color_manual(labels = c("k=3", "k=4", "k=5", "k=6"),
                     values = c("red", "green", "blue", "darkred"))
```

